---
title: "Metadata Labling Template"
author: "Jane Velghe"
date: '2022-08-03'
output: pdf_document
---

```{r }
library(dplyr)
library(Seurat)
library(ggplot2)
library(patchwork)
library(DropletUtils)
# We will repeat this process for each aggregated output of the samples 
# Initialize the Seurat object with the raw (non-normalized data).
# We start by reading in the data. The Read10X() function reads in the output
# of the cellranger pipeline from 10X, returning a unique molecular 
# identified (UMI) count matrix. The values in this matrix represent the 
# number of molecules for each feature (i.e. gene; row) that are detected 
# in each cell (column).
```

What's the name of the folder for this sample? - all you need to change!
```{r }
name <- '/R441P' #change
folder <- '/R441' #change

sample <- 'R441P' #change
donor <- 'R441' #change
```

Update metadata variables
```{r }
cells <- 'CD45_Positive' 
#cells <- 'CD45_Negative' 
donor_sample <- paste0(donor,'_Immune')
#donor_sample <- paste0(donor,'_Islet')
```

Let's get organized before we start
```{r }
#path to project folder
data_path <- c('~/Verchere/jdrf2')

#make a folder for objects plots and name a filepath for this
dir.create(paste0(data_path,folder,name,"/filt_outs"))
filt_outs <- paste0(data_path,folder,name,"/filt_outs")
#make a folder for plots before we filter and name a filepath for this
dir.create(paste0(data_path,folder,name,"/prefilt_outs"))
prefilt_outs <- paste0(data_path,folder,name,"/prefilt_outs")
#make a folder for doubletfinder plots and name a filepath for this
dir.create(paste0(data_path,folder,name,"/objects"))
objects <- paste0(data_path,folder,name,"/objects")
```

```{r}
data_dir <- paste0(data_path,folder,name,'/SoupX/soupX_filt')
list.files(data_dir) # Should show barcodes.tsv.gz, features.tsv.gz, and matrix.mtx.gz
data <- Read10X(data.dir = data_dir)
obj_prefilt <- CreateSeuratObject(counts = data, project = sample, min.cells = 3, min.features = 200) #min.cells = 3, min.features = 200
```

```{r }
#Add labels
levels(obj_prefilt) #check what identity class labels it has from the start are
Idents(obj_prefilt) <- "orig.ident" #let's call this identity class this
#now let's add more metadata to this sample dataset

cells.use <- WhichCells(object = obj_prefilt, idents = sample) #WHO -> which cells do we want to label
obj_prefilt <- SetIdent(object = obj_prefilt, cells = cells.use, value = sample) #WHAT -> what do we we want to call them?
obj_prefilt[["Sample"]] <- Idents(object = obj_prefilt) #Where -> Where are we storing this new identity? 
# here we make a new identity class called sample type, this sample contains PBMCs. 
levels(obj_prefilt)
# identity now has changed to most recent value

cells.use <- WhichCells(object = obj_prefilt, idents = sample) #WHO -> which cells do we want to label
obj_prefilt <- SetIdent(object = obj_prefilt, cells = cells.use, value = donor) #WHAT -> what do we we want to call them?
obj_prefilt[["Donor"]] <- Idents(object = obj_prefilt) #Where -> Where are we storing this new identity? 
# here we make a new identity class called sample type, this sample contains PBMCs. 
levels(obj_prefilt)
# identity now has changed to most recent value

cells.use <- WhichCells(object = obj_prefilt, idents = donor) #WHO -> which cells do we want to label
obj_prefilt <- SetIdent(object = obj_prefilt, cells = cells.use, value = cells) #WHAT -> what do we we want to call them?
obj_prefilt[["Cells"]] <- Idents(object = obj_prefilt) #Where -> Where are we storing this new identity? 
# here we make a new identity class called sample type, this sample contains PBMCs. 
levels(obj_prefilt)
# identity now has changed to most recent value

#let's make another identity, this time it will just be of the machine used by 10X on these samples
cells.use <- WhichCells(object = obj_prefilt, idents = cells) #you can only get idents from those in a class that is active
obj_prefilt <- SetIdent(object = obj_prefilt, cells = cells.use, value = donor_sample)
obj_prefilt[["Donor_Cells"]] <- Idents(object = obj_prefilt)
levels(obj_prefilt)

#switch back to the original identity
Idents(obj_prefilt) <- "orig.ident"
levels(obj_prefilt) 
```

```{r}
# Add cell metadata
obj_prefilt[["percent.mt"]] <- PercentageFeatureSet(obj_prefilt, pattern = "MT-")
obj_prefilt[["percent.rb"]] <- PercentageFeatureSet(obj_prefilt, pattern = "RP-")
obj_prefilt <- CellCycleScoring(
  object = obj_prefilt,
  g2m.features = cc.genes$g2m.genes,
  s.features = cc.genes$s.genes,
  nbin = 20)
```

Save unfiltered seurat object as a .rds file
```{r}
saveRDS(obj_prefilt, paste0(objects,name,"_pre-filtered_so.rds"))
```


Visualize QC metrics as a violin plot
```{r}
VlnPlot(obj_prefilt, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), pt.size = 0.1, assay ="RNA", ncol = 4)
ggsave(paste0(prefilt_outs,name,"_vln_raw_point_plot_by_run.pdf"), plot = last_plot(), height = 15, width = 10)
```

Make scatterplot 
```{r}
FeatureScatter(obj_prefilt, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
ggsave(paste0(prefilt_outs,name,"_scatter_plot.pdf"), plot = last_plot(), height = 10, width = 10)
```

Check cycling cells
```{r}
VlnPlot(obj_prefilt, features = c("S.Score","G2M.Score"), pt.size =  0.01)
ggsave(paste0(prefilt_outs,name,"_cell_cycle_vln_plot.pdf"), plot = last_plot(), height = 10, width = 10)
```

Remove bad cells -> this is common practice. 
```{r}
obj_filt <- obj_prefilt

#remove cells with nFeature_RNA < 7000
#obj_filt <- subset(obj_filt, subset = nFeature_RNA < 7000) 

#remove cells with MT% >10%
obj_filt <- subset(obj_filt, subset = percent.mt < 10) 

#remove cycling cells 
#obj_filt <- subset(obj_filt, subset = G2M.Score < 0.5) 
#obj_filt <- subset(obj_filt, subset = S.Score < 0.5) 
```

Visualize QC metrics as a violin plot
```{r}
VlnPlot(obj_filt, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.rb"), pt.size = 0.1, assay ="RNA", ncol = 4)
ggsave(paste0(filt_outs,name,"_vln_raw_point_plot_by_run.pdf"), plot = last_plot(), height = 15, width = 10)
```

```{r}
FeatureScatter(obj_filt, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
ggsave(paste0(filt_outs,name,"_scatter_plot.pdf"), plot = last_plot(), height = 10, width = 10)
```

```{r}
VlnPlot(obj_filt, features = c("S.Score","G2M.Score"), pt.size =  0.01)
ggsave(paste0(filt_outs,name,"_cell_cycle_vln_plot.pdf"), plot = last_plot(), height = 10, width = 10)
```

Save filtered seurat object as a .rds file
```{r}
saveRDS(obj_filt, paste0(objects,name,"_filtered_so.rds"))
```